Identification of high-yielding soybean lines with exceptional seed composition qualities

### 2022-06-05 doc/manuscript/manuscript.Rmd
Added to manuscript

```{r setup, warning=FALSE,echo = FALSE, message=FALSE}
library(here)
source(here("packages.R"))
tar_load(c(example_table, 
           yield_summary_tables, 
           elite_genotype_tables, 
           phenotype_summary_plots))
```

@@

Eighty unique lines were selected from each population which satisfied these criteria, and each yield test was comprised of 40 RILs. Three high-yielding check cultivars and the two parents of the respective population were also included in each test. Yield check cultivars Dunphy, Osage[@Osage], and Roy were used in tests 1 and 2, while Dunphy, Dilday, and NC-Raleigh[@NCRaleighregistration] were used for tests 3 and 4. These lines were selected to represent the estimated maturities of the RILs in each test. The parents for tests 1 and 2 were cultivars LMN09-119 and N09-09, and the parents for tests 3 and 4 were LMN09-19 and N13-47.

@@

These thirty lines were then grouped into two new tests of 15 RILs each based on maturity date. These two new tests are named Test 1 and Test 2. Yield check cultivars were again assigned to each test to match the maturity dates of the RILs that were in each test. Cultivars Dunphy, DIlday, and NC-Raleigh were used as checks in Test 1 and Dunphy, Ellis, N10-697, and Osage were used as checks in Test 2.

@@

## Yield contrasts
Several RILs in each of the tests had yield that was comparable to that of their check cultivars. This was determined both quantitatively through the use of contrasts between the mean of each RIL with the average of the yield check means, and qualitatively by inspecting the standard errors on the estimation of the genotype marginal means for yield. No genotypes had yield that was quantitatively higher than that of the yield checks, however many had comparable yield. Many of these genotypes with comparable yield also had protein content that was superior to the check average in each test. No genotypes had oil that was superior to the check cultivars, but some had comparable yield and oil content as well as superior protein content.

@@

A strong negative correlation was observed in both populations betweeen seed oil and seed protein. This observation matches well with previous findings on the inverse correlation between seed protein and seed oil. Many RILs had high protein when compared with the check cultivars so it was not surprising that relatively few had both a superior seed protein content, and a comparable seed oil content.

## Genotypes with comparable yield and superior protein
Seventeen RILs had a similar yield and superior protein content to the check cultivars. Summary data for the seed protein and seed yield of these lines is given in table 2. The maximum yielding line

@@

Among these genotypes are several with both average seed protein and average seed yield that were greater than the averages of the checks for each test. These N18-1632-1, N18-1620, N18-1635 from test 1 and RILs N18-1761 and N18-1627 from test 2. Of particular note are genotypes N18-1632-1 from test 1 and N18-1761 from test 2. On average, genotype N18-1632-1 had a protein content that was 113.9% of the check average while maintaining a yield average that was 101.5% that of the check average.

>>>>

```{r setup, warning=FALSE,echo = FALSE, message=FALSE}
library(here)
source(here("packages.R"))
tar_load(c(yield_summary_tables, 
           elite_genotype_tables, 
           phenotype_summary_plots))
```

@@

Eighty unique lines were selected from each population which satisfied these criteria, and each yield test was comprised of 40 RILs. Three high-yielding check cultivars and the two parents of the respective population were also included in each test. These yield tests were named test 1 and test 2 for RILs derived from mapping population 201 and test 3 and 4 for RILs derived from mapping population 202. Yield check cultivars Dunphy, Osage[@Osage], and Roy were used in tests 1 and 2, while Dunphy, Dilday, and NC-Raleigh[@NCRaleighregistration] were used for tests 3 and 4. These lines were selected to represent the estimated maturities of the RILs in each test. The parents for tests 1 and 2 were cultivars LMN09-119 and N09-09, and the parents for tests 3 and 4 were LMN09-19 and N13-47.

@@

These thirty lines were then grouped into two new tests of 15 RILs each based on maturity date. These two new tests are named Test 1 and Test 2. Yield check cultivars were again assigned to each test to match the maturity dates of the RILs that were in each test. Cultivars Dunphy, Dilday, and NC-Raleigh were used as checks in Test 1 and Dunphy, Ellis, N10-697, and Osage were used as checks in Test 2.

@@

Genotype least square means were calculated for all RILs in both tests. A visualization of these least square means and the standard error on the estimation of these means for seed protein, seed oil, and seed yield can be seen in figure 1. Both tests tended to have more RILs with high protein rather than high oil when compared to the average of the checks, and a moderate number had a yield average greater than the average of the checks. Qualitatively, many RILs had a comparable yield to the checks following the relatively large standard error associated with the estimating the yield, and a qualitatively larger seed protein content given the relatively small standard error associated with the estimation of seed protein content. To separate means quantitatively, we performed contrasts between each RIL mean and the average of the checks for each test.

## Yield contrasts
Several RILs in each of the tests had yield that was comparable to that of their check cultivars as per contrasts between RIL means and the average of the check cultivars for each test. No genotypes had yield that was quantitatively higher than that of the yield checks, however many had comparable yield. Many of these genotypes with comparable yield also had protein content that was superior to the check average in each test. No genotypes had oil that was superior to the check cultivars, but some had comparable yield and oil content as well as superior protein content.

@@

A strong negative correlation was observed in both populations between seed oil and seed protein, and a weak inverse correlation was observed between seed protein and seed yield. This observation matches well with previous findings on the inverse correlation between seed protein and seed oil, and between seed protein and seed yield. Many RILs had high protein when compared with the check cultivars so it was not surprising that relatively few had both a superior seed protein content, and a comparable seed oil content.

These RILs can provide valuable germplasm that can be used to simultaneously improve both seed oil and seed protein content without compromising seed yield or agronomics. This can be of use both to breeders that are looking to improve seed composition traits in soybean and for growers who are seeking options for novel soybean varieties with seed composition that is superior to that of cultivars that are already used in production.

## Genotypes with comparable yield and superior protein
Seventeen RILs had a similar yield and superior protein content to the check cultivars. Summary data for the seed protein and seed yield of these lines is given in table 2. From among these RILs, N18-1627 had the highest seed yield relative to the checks included in the test with an average yield that was 105.9% that of the check average of Test 2. Genotype N18-1763 had the highest seed protein content on average with a protein content that was 116.8% that of the check average of Test 1. We observed a negative correlation between seed protein and seed yield in both tests so RILs with seed protein content that was above the check average tended to have a seed yield content that was comparatively lower.

@@

However, among these genotypes are several with both average seed protein and average seed yield that were greater than the averages of the checks for each test. These N18-1632-1, N18-1620, N18-1635 from test 1 and RILs N18-1761 and N18-1627 from test 2. Of particular note are genotypes N18-1632-1 from test 1 and N18-1761 from test 2. These two RILs have a seed protein content that is substantially above that of the average protein content of he checks while also maintaining an average yield that is above the average yield of the checks. These RILs are ideal candidates for use in breeding programs which seek to improve seed protein content in particular without compromising seed yield or argonomics.

### 2022-06-05 doc/manuscript/manuscript.tex
Added to manuscript

Eighty unique lines were selected from each population which satisfied
these criteria, and each yield test was comprised of 40 RILs. Three
high-yielding check cultivars and the two parents of the respective
population were also included in each test. Yield check cultivars
Dunphy, Osage\citep{Osage}, and Roy were used in tests 1 and 2, while
Dunphy, Dilday, and NC-Raleigh\citep{NCRaleighregistration} were used
for tests 3 and 4. These lines were selected to represent the estimated
maturities of the RILs in each test. The parents for tests 1 and 2 were
cultivars LMN09-119 and N09-09, and the parents for tests 3 and 4 were
LMN09-19 and N13-47.

@@

\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{C:/Users/Jay/Documents/R/Projects/Jay_yield/exports/plots/lsmean_scatterplot.pdf}
\caption{Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (\textbf{A}) and Test 2 (\textbf{B}). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.}
\end{figure}

@@

Several RILs in each of the tests had yield that was comparable to that
of their check cultivars. This was determined both quantitatively
through the use of contrasts between the mean of each RIL with the
average of the yield check means, and qualitatively by inspecting the
standard errors on the estimation of the genotype marginal means for
yield. No genotypes had yield that was quantitatively higher than that
of the yield checks, however many had comparable yield. Many of these
genotypes with comparable yield also had protein content that was
superior to the check average in each test. No genotypes had oil that
was superior to the check cultivars, but some had comparable yield and
oil content as well as superior protein content.

@@

A strong negative correlation was observed in both populations betweeen
seed oil and seed protein. This observation matches well with previous
findings on the inverse correlation between seed protein and seed oil.
Many RILs had high protein when compared with the check cultivars so it
was not surprising that relatively few had both a superior seed protein
content, and a comparable seed oil content.

@@

Seventeen RILs had a similar yield and superior protein content to the
check cultivars. Summary data for the seed protein and seed yield of
these lines is given in table 2. The maximum yielding line

@@

Among these genotypes are several with both average seed protein and
average seed yield that were greater than the averages of the checks for
each test. These N18-1632-1, N18-1620, N18-1635 from test 1 and RILs
N18-1761 and N18-1627 from test 2. Of particular note are genotypes
N18-1632-1 from test 1 and N18-1761 from test 2. On average, genotype
N18-1632-1 had a protein content that was 113.9\% of the check average
while maintaining a yield average that was 101.5\% that of the check
average.

>>>>

Eighty unique lines were selected from each population which satisfied
these criteria, and each yield test was comprised of 40 RILs. Three
high-yielding check cultivars and the two parents of the respective
population were also included in each test. These yield tests were named
test 1 and test 2 for RILs derived from mapping population 201 and test
3 and 4 for RILs derived from mapping population 202. Yield check
cultivars Dunphy, Osage\citep{Osage}, and Roy were used in tests 1 and
2, while Dunphy, Dilday, and NC-Raleigh\citep{NCRaleighregistration}
were used for tests 3 and 4. These lines were selected to represent the
estimated maturities of the RILs in each test. The parents for tests 1
and 2 were cultivars LMN09-119 and N09-09, and the parents for tests 3
and 4 were LMN09-19 and N13-47.

@@

\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{C:/Users/Jay/Desktop/Documents/R/Jay_yield/exports/plots/lsmean_scatterplot.pdf}
\caption{Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (\textbf{A}) and Test 2 (\textbf{B}). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.}
\end{figure}

Genotype least square means were calculated for all RILs in both tests.
A visualization of these least square means and the standard error on
the estimation of these means for seed protein, seed oil, and seed yield
can be seen in figure 1. Both tests tended to have more RILs with high
protein rather than high oil when compared to the average of the checks,
and a moderate number had a yield average greater than the average of
the checks. Qualitatively, many RILs had a comparable yield to the
checks following the relatively large standard error associated with the
estimating the yield, and a qualitatively larger seed protein content
given the relatively small standard error associated with the estimation
of seed protein content. To separate means quantitatively, we performed
contrasts between each RIL mean and the average of the checks for each
test.

@@

Several RILs in each of the tests had yield that was comparable to that
of their check cultivars as per contrasts between RIL means and the
average of the check cultivars for each test. No genotypes had yield
that was quantitatively higher than that of the yield checks, however
many had comparable yield. Many of these genotypes with comparable yield
also had protein content that was superior to the check average in each
test. No genotypes had oil that was superior to the check cultivars, but
some had comparable yield and oil content as well as superior protein
content.

@@

A strong negative correlation was observed in both populations between
seed oil and seed protein, and a weak inverse correlation was observed
between seed protein and seed yield. This observation matches well with
previous findings on the inverse correlation between seed protein and
seed oil, and between seed protein and seed yield. Many RILs had high
protein when compared with the check cultivars so it was not surprising
that relatively few had both a superior seed protein content, and a
comparable seed oil content.

These RILs can provide valuable germplasm that can be used to
simultaneously improve both seed oil and seed protein content without
compromising seed yield or agronomics. This can be of use both to
breeders that are looking to improve seed composition traits in soybean
and for growers who are seeking options for novel soybean varieties with
seed composition that is superior to that of cultivars that are already
used in production.

@@

Seventeen RILs had a similar yield and superior protein content to the
check cultivars. Summary data for the seed protein and seed yield of
these lines is given in table 2. From among these RILs, N18-1627 had the
highest seed yield relative to the checks included in the test with an
average yield that was 105.9\% that of the check average of Test 2.
Genotype N18-1763 had the highest seed protein content on average with a
protein content that was 116.8\% that of the check average of Test 1. We
observed a negative correlation between seed protein and seed yield in
both tests so RILs with seed protein content that was above the check
average tended to have a seed yield content that was comparatively
lower.

@@

However, among these genotypes are several with both average seed
protein and average seed yield that were greater than the averages of
the checks for each test. These N18-1632-1, N18-1620, N18-1635 from test
1 and RILs N18-1761 and N18-1627 from test 2. Of particular note are
genotypes N18-1632-1 from test 1 and N18-1761 from test 2. These two
RILs have a seed protein content that is substantially above that of the
average protein content of he checks while also maintaining an average
yield that is above the average yield of the checks. These RILs are
ideal candidates for use in breeding programs which seek to improve seed
protein content in particular without compromising seed yield or
argonomics.
